---
title: "Final Project DS4EEB"
author: "Hannah Hayes"
format: html
editor: visual
---

## Data Wrangling

### Libraries & Packages

```{r}
#install.packages("usethis")
library("usethis")
#install.packages("here")
library("here")
#install.packages("dplyr")
library(tidyverse) #this will help us clean up the data 
library(readr)
#install.packages("respR")
library(respR)
#install.packages("devtools")
library(devtools)
library(lubridate) #To turn time HH:MM:SS AM/PM into seconds passed
#install.packages("rTPC")
library(rTPC)
library(dplyr)
library(ggplot2)
library(brms)
```

### Working Directory

```{r}
#set the working directory
setwd("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry") 
```

### Read in the Data

I am having a lot of trouble reading the data back in. When I was working through the data wrangling I was periodically saving the new CSVs. Chunks 3 through 8 will not run. The CSVs from the UTF folder are no longer acceptable csv format. I'm not sure why. The files need to be converted back un UTF-8 files. I tried to do the read.csv(..., UTF file...). It did not work.

```{r}
#| eval: false
#| include: false


#Field Resp Chamber 1 
file.choose()
FRCh1 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber1_218202610216_UTF.csv")
FRCh1 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh1 <- FRCh1 %>%
  select(-Date) %>% 
 mutate(chamber = 1, Date = "2026-02-18")

#Field Resp Chamber 2 
file.choose()
FRCh2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber2_218202610216_UTF.csv")
FRCh2 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh2 <- FRCh2 %>%
  select(-Date) %>% 
 mutate(chamber = 2, Date = "2026-02-18")

#Field Resp Chamber 3
FRCh3 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber3_218202610216_UTF.csv")
FRCh3 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh3 <- FRCh3 %>%
 mutate(chamber = 3, Date = "2026-02-18")

#Field Resp Chamber 4
FRCh4 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber4_218202610216_UTF.csv")
FRCh4 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh4 <- FRCh4 %>%
 mutate(chamber = 4, Date = "2026-02-18")

#Field Resp Chamber 5
FRCh5 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber5_218202610216_UTF.csv")
FRCh5 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh5 <- FRCh5 %>%
 mutate(chamber = 5, Date = "2026-02-18")

#Field Resp Chamber 6
FRCh6 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber6_218202610216_UTF.csv")
FRCh6 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh6 <- FRCh6 %>%
 mutate(chamber = 6, Date = "2026-02-18")

#Field Resp Chamber 7
FRCh7 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber7_218202610216_UTF.csv")
FRCh7 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh7 <- FRCh7 %>%
 mutate(chamber = 7, Date = "2026-02-18")

#Field Resp Chamber 8
FRCh8 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber8_218202610216_UTF.csv")
FRCh8 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh8 <- FRCh8 %>%
 mutate(chamber = 8, Date = "2026-02-18")

```

### Combine all of the CSVs

Must combine all of the CSVs into one

```{r}
#| eval: false
#| include: false
FR <- bind_rows(FRCh1, FRCh2, FRCh3, FRCh4, FRCh5, FRCh6, FRCh7, FRCh8) #Combine all of the data to one CSV
FR #Check that it worked

unique(FR$chamber) #Check that all the chambers made it into the same CSV
```

### Convert AM/PM into 24hr

The time is the format of HH:MM:SS AM/PM format and needs to converted into HH:MM:SS

```{r}
#| eval: false
#| include: false
# Call upon Time column 
FR$Time

# Convert to a time object with a temporary date, recognizing the AM/PM format
# strptime() automatically converts to 24-hour format internally
time_objects <- strptime(FR$Time, format = "%I:%M:%S %p")

# Format the time objects to display only the time in 24-hour format (military time)
FR$Time <- format(time_objects, format = "%H:%M:%S")

# Print the result
FR$Time
```

### Convert Time into Seconds

Then time needs to be broken down in to seconds passed. Which from each time step should be one second.

```{r}
#| eval: false
#| include: false
FR_sec <- FR %>%
  mutate(
    time_parsed = hms(Time),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Select for columns

```{r}
#| eval: false
#| include: false
#O2%, Chamber, Date, seconds, temperature
# Select specific columns
FR_resp <- select(FR_sec, chamber, Date, )
# Or use the pipe operator (%>%) which is common in Tidyverse
FR_resp <- FR_sec %>% 
  select(column_name_1, column_name_2)
FR_resp$colname
```

### Change Time into seconds

```{r}
#| eval: false
#| include: false
FR_sec <- FR %>%
  mutate(
    time_parsed = parse_date_time(Time, "I:M:S"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )

# Source - https://stackoverflow.com/a/31549819
# Posted by inhuretnakht, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-06, License - CC BY-SA 4.0

library(lubridate)
FR_sec <- FR %>% 
period_to_seconds(hms("10:02:17"))

FR_sec <- FR %>%
  filter(is.na(parse_date_time(Time, orders = "I:M:S p"))) %>%
  head()

FR_Time <- trimws(FR, which = Time)

FR_sec <- FR_Time %>%
  mutate(
    time_parsed = parse_date_time(Time, orders = "I:M:S p"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Duplicate NA Columns.

The column names were not exactly the same so it duplicated the columns and had a bunch of NAs.

```{r}
#FR_sec
#write.csv(FR_sec, "FR_sec.csv")
#file.choose()
FR_2 <- read_csv("Data/FieldRep_021826-2/UTF/FR.csv")
#FR_2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FR.csv")

Chamber1 <- FR_2 %>% #we need to get rid of salinity -> TempComp
  filter(chamber == 1) %>% # we only want chamber 1 rows
  select(-(Salinity:TempComp)) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Time, Intermittent_phase, O_perecent_airsat:Temperature_compensation_C, chamber:seconds)


Chamber2_8 <- FR_2 %>% 
  filter(chamber > 1) %>% 
  select(-(Salinity_N:Temperature_compensation_C)) %>% 
  relocate(Salinity:TempComp, .before = PowerX4_EAC7_voltage_V_flush) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Time, Intermittent_phase, Oairsat:TempComp, chamber:seconds)

colnames(Chamber1) <- colnames(Chamber2_8) #renaming all of the columns in the Chamber1 dataframe with the columns names from Chamber2_8
FieldResp <- rbind(Chamber1, Chamber2_8) #Combining the two dataframes as FieldResp
write.csv(FieldResp, "FieldResp_2026-02-18.csv")

FieldResp <- read.csv("FieldResp_2026-02-18.csv") #save into a new csv

FieldResp %>% 
  ggplot(aes(x = seconds, y = Oairsat, color = chamber)) +
  geom_point()


Field_Resp_select <- FieldResp %>% 
  mutate(Tub = ifelse(chamber <= 4, "A", "B")) %>% 
  select(Time, Intermittent_phase, Oairsat,Temp, chamber, Date, seconds, Tub) 

```

### Calculate oxygen consumed for every time step

```{r}
#we're working with the field resp select data frame because I need the loop column 
head(Field_Resp_select)

 # we need these columns timepassed = 1, Oairsar2, tempK, metabolic rate, blank rate, o2seccorrected, o2 mg hour
Field_Reskp_grouped <- Field_Resp_select %>%
  arrange(Tub, chamber, Time) %>%   # sort first by blank,chamber, then time
  group_by(Tub, chamber) %>%        # calculate rates within each replicate
  mutate(umol.sec = c(NA, diff(Oairsat) / diff(seconds))) #%>% 
 # ungroup() %>%
 # mutate(blank.rate = ifelse(chamber == 4, umol.sec, NA)) 
```

### Blanks

When doing respirometry, you typically run a blank chamber that does not have an organism in it to account for any background noise (respiration, photosynthesis) that may be affecting the O2 measurements that cannot be attributed to your study organism. We must calculate the oxygen consumption rates of the blanks and minus them from the actual rates from the other chambers.

```{r}
#Now we need the blank rate to minus 
#Get the blank rates 
blank1_rate <- Field_Reskp_grouped %>%
  filter(chamber == 4) %>%
  pull(umol.sec)
  

blank2_rate <- Field_Reskp_grouped %>%
  filter(chamber == 8) %>%
  pull(umol.sec)

#Create the blank_rate column 
Field_Reskp_adj <- Field_Reskp_grouped %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ blank1_rate,
      chamber %in% 5:7 ~ blank2_rate,
      TRUE ~ NA_real_
    )
  )%>% 
  filter(!chamber == "4" | !chamber == "8") 

```

### Group by temperature bin

```{r}
#seperate m from numbers in the intermittent phase column
Field_Reskp_adj <- Field_Reskp_adj %>%
  mutate(Intermittent_phase = as.numeric(gsub("M", "", Intermittent_phase)))

#Lets add in a temp bin
Field_Reskp_tbin <- Field_Reskp_adj %>% 
mutate(temp_bin = round(Temp))
```

### Aggregate Function

Grouping are not giving us a mean with in the correct order. Let's try it using the aggregate function!

```{r}
#We're going to group it differently this time
aggy <- aggregate(umol.sec ~ Tub*temp_bin*chamber, data=Field_Reskp_tbin, mean)


#minus the blanks 
blank1_rate <- aggy %>%
  slice(-148) %>%
  filter(chamber == 4) %>%
  pull(umol.sec)
  

blank2_rate <- aggy %>%
  filter(chamber == 8) %>%
  pull(umol.sec)

#Create the blank_rate column 
aggy_adj <- aggy %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ first(blank1_rate),
      chamber %in% 5:7 ~ first(blank2_rate),
      TRUE ~ NA_real_
    )
  )#%>% 
  #filter(!chamber == "4" | !chamber == "8") 

#Minus the blank rate and the umoll to get the adjusted value
aggy_corr <- aggy_adj %>% 
  mutate(O2.sec.corr = umol.sec - blank_rate) 

#Correct for the hour, but we need the mass
aggy_hh <- aggy_corr %>% 
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/1000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/1000
    )
  ) %>% 
  mutate(umol_cm2_hr = (O2.sec.corr * 3600) / masskg)

#remove NAs
aggy_na <- aggy_hh %>% 
    filter(!is.na(umol_cm2_hr))
#make it positive
  aggy_pos <- aggy_na %>% 
    mutate(umol_cm2_hr = -(umol_cm2_hr)) %>% 
    filter(umol_cm2_hr > 0)
  
#log it for get the rate
   aggy_log <- aggy_pos %>% 
    mutate(rate = log(umol_cm2_hr + 0.01))
   
#we fianlly got some good data!
write.csv(aggy_log, "TPCdata.csv")
```

### Lets check the visuals

```{r}
#logged Rate vs temp bin 
plot(aggy_log$temp_bin,
     aggy_log$rate,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")

plot(aggy_log$temp_bin,
     aggy_log$umol_cm2_hr,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")
```

Really cool! So It's giving us a general TPC, but there is a decent amount of noise and not a clear drop off. I believe we need to ramp the urchins up to a higher temperature to get the end of the TPC.

## Analysis

### Read in the data

```{r}
#read in the data
TPCdata <- read.csv("TPCdata.csv")
```

### Let's visualize the data

```{r}

model <- TPCdata %>% 
  ggplot(aes(x = temp_bin, y = rate)) +
  geom_point() 
model
```

### Is it normal data?

```{r}
hist <- TPCdata %>% 
  ggplot(aes(x = rate)) +
  geom_histogram()
hist

hist2 <- TPCdata %>% 
  ggplot(aes(x = umol_cm2_hr)) +
  geom_histogram()
hist2
```

### BRMS

```{r}
fit_simple <- 
  brm(data = TPCdata, # Give the model the data
      # gaussian (normal) distribution
      family = gaussian,
      rate ~ 1 + temp_bin,
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 4,
      file = "output/fit_simple")

summary(fit_simple)
plot(fit_simple)
```

### I noticed in the histogram the data looks a bit skewed so I would like to try using a gamma family instead of a Guassain

```{r}
fit_gamma <- 
  brm(data = TPCdata, # Give the model the data
      # Choose a gaussian (normal) distribution
      family = Gamma(),
      # Specify the model here. 
      rate ~ 1 + temp_bin,
      # Here's where you specify parameters for executing the Markov chains
      # We're using similar to the defaults, except we set cores to 4 so the analysis runs faster than the default of 1
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      # Setting the "seed" determines which random numbers will get sampled.
      # In this case, it makes the randomness of the Markov chain runs reproducible 
      # (so that both of us get the exact same results when running the model)
      seed = 4,
      # Save the fitted model object as output - helpful for reloading in the output later
      file = "output/fit_gamma")

summary(fit_gamma)
plot(fit_gamma)
```
