---
title: "Final Project DS4EEB"
author: "Hannah Hayes"
format: html
editor: visual
---

### Libraries & Packages

```{r}
#install.packages("usethis")
library("usethis")
#install.packages("here")
library("here")
#install.packages("dplyr")
library(tidyverse) #this will help us clean up the data 
library(readr)
#install.packages("respR")
library(respR)
#install.packages("devtools")
library(devtools)
library(lubridate) #To turn time HH:MM:SS AM/PM into seconds passed
```

### Working Directory

```{r}
#file.choose()
setwd("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF") 
```

### Read in the Data

The way that LoLigo has give us the data is a separate CSV for each chamber, and they don't have a column for which chamber it is. So I'm going to read in all the data and add a column for the chamber number.

```{r}
#Field Resp Chamber 1 
FRCh1 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber1_218202610216_UTF.csv")
FRCh1 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh1 <- FRCh1 %>%
  select(-Date) %>% 
 mutate(chamber = 1, Date = "2026-02-18")

#Field Resp Chamber 2 
FRCh2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber2_218202610216_UTF.csv")
FRCh2 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh2 <- FRCh2 %>%
  select(-Date) %>% 
 mutate(chamber = 2, Date = "2026-02-18")

#Field Resp Chamber 3
FRCh3 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber3_218202610216_UTF.csv")
FRCh3 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh3 <- FRCh3 %>%
 mutate(chamber = 3, Date = "2026-02-18")

#Field Resp Chamber 4
FRCh4 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber4_218202610216_UTF.csv")
FRCh4 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh4 <- FRCh4 %>%
 mutate(chamber = 4, Date = "2026-02-18")

#Field Resp Chamber 5
FRCh5 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber5_218202610216_UTF.csv")
FRCh5 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh5 <- FRCh5 %>%
 mutate(chamber = 5, Date = "2026-02-18")

#Field Resp Chamber 6
FRCh6 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber6_218202610216_UTF.csv")
FRCh6 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh6 <- FRCh6 %>%
 mutate(chamber = 6, Date = "2026-02-18")

#Field Resp Chamber 7
FRCh7 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber7_218202610216_UTF.csv")
FRCh7 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh7 <- FRCh7 %>%
 mutate(chamber = 7, Date = "2026-02-18")

#Field Resp Chamber 8
FRCh8 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber8_218202610216_UTF.csv")
FRCh8 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh8 <- FRCh8 %>%
 mutate(chamber = 8, Date = "2026-02-18")

```

### Combine all of the CSVs

```{r}
FR <- bind_rows(FRCh1, FRCh2, FRCh3, FRCh4, FRCh5, FRCh6, FRCh7, FRCh8) #Combine all of the data to one CSV
FR #Check that it worked

unique(FR$chamber) #Check that all the chambers made it into the same CSV
```

### For Loop

```{r}

```

### Convert AM/PM into 24hr

```{r}
# Call upon Time column 
FR$Time

# Convert to a time object with a temporary date, recognizing the AM/PM format
# strptime() automatically converts to 24-hour format internally
time_objects <- strptime(FR$Time, format = "%I:%M:%S %p")

# Format the time objects to display only the time in 24-hour format (military time)
FR$Time <- format(time_objects, format = "%H:%M:%S")

# Print the result
FR$Time
```

### Convert Time into Seconds

```{r}
FR_sec <- FR %>%
  mutate(
    time_parsed = hms(Time),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Select for variables

```{r}
#O2%, Chamber, Date, seconds, temperature
# Select specific columns
FR_resp <- select(FR_sec, chamber, Date, )
# Or use the pipe operator (%>%) which is common in Tidyverse
FR_resp <- FR_sec %>% 
  select(column_name_1, column_name_2)
FR_resp$colname
```

### Trials

#### Trial #1

okay so we're having way more trouble loading in the data. I think we may have got it now so let's keep going. So through a lot of trial and error and googling and using chatgpt. Turns out the orgiianl file was a UTF 16, but r uses a UTF -8 files. So using TextEdit I was able to change the type of file then i reopened it in ecel and turned it into a CSBV now it works!!!!!!

#### Trial #2

Don't forget to check that everything is there. I didn't notice till after all of these steps that the seconds were missing from the time column so I have to go back and fix all of the csv files again!!!!

```         
```

### Snips

#### 1st attempt at bindrows

This was to bind all of the CSVs from my local file together, but then I forgot none of them actually have the chamber column

`{r| eval: false} #FR <- list.files(   #path = "/Users/hannahhayes858/Desktop/JJG x HH Collab/CSUMB-UCSC-Urchin-Respirometry/FieldResp_021826/UT8s/TestingforUTF",   #full.names = TRUE #) %>%   # lapply(read_csv) %>%   # bind_rows() #FR #write.csv(FR, "fieldrespch1_8.csv")`

#### Separate Date and Time

\`\`\`{r\| eval: false} #First we need to seperate the date and the time

df_char \<- data.frame( datetime_string = c("2021-05-25 12:34:25", "2019-01-14 04:10:30"))

df_separated \<- FR %\>% \# separate(Datetime, into = c("date", "time"), sep = " ") df_separated \`\`\`

#### Sardine Data Set Example

\`\`\`{r\| eval: false} #\| eval: false #\| include: false \# Example using built-in data 'sardine.rd' data(sardine.rd) \# Inspect the data (time in column 1, oxygen in column 2) sardine_insp \<- inspect(sardine.rd, time = 1, oxygen = 2)

Calculate rate from a specific region (manual)

manual_rate \<- calc_rate(sardine_insp, from = 1000, to = 3000)

Automatically detect the "best-fit" linear rate

auto_rate_obj \<- auto_rate(sardine_insp, method = "linear")

Calculate background rate from background data (example data not shown)

bg_rate \<- calc_rate.bg(background_data)

Adjust the specimen rate

adjusted_rate \<- adjust_rate(auto_rate_obj, by = bg_rate)

Convert to mass-specific units (requires specimen mass and chamber volume)

final_rate \<- convert_rate(adjusted_rate, volume = 1.2, \# volume of respirometer in L mass = 0.5, \# mass of specimen in kg \# input/output units will depend on your data O2.unit = "mg/L", time.unit = "hours", output.unit = "mg/h/kg")

```{r}
#devtools::install_github("januarharianto/respR")


#load the data and combine all csv files
# Source - https://stackoverflow.com/a/30242859
# Posted by Maiasaura, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-04, License - CC BY-SA 4.0
```

#### Change Time into seconds

```{r}
#| eval: false
#| include: false
FR_sec <- FR %>%
  mutate(
    time_parsed = parse_date_time(Time, "I:M:S"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )

# Source - https://stackoverflow.com/a/31549819
# Posted by inhuretnakht, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-06, License - CC BY-SA 4.0

library(lubridate)
FR_sec <- FR %>% 
period_to_seconds(hms("10:02:17"))

FR_sec <- FR %>%
  filter(is.na(parse_date_time(Time, orders = "I:M:S p"))) %>%
  head()

FR_Time <- trimws(FR, which = Time)

FR_sec <- FR_Time %>%
  mutate(
    time_parsed = parse_date_time(Time, orders = "I:M:S p"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Duplicate NA Columns

```{r}
FR_sec
write.csv(FR_sec, "FR_sec.csv")
file.choose()
FR_2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FR.csv")

Chamber1 <- FR_2 %>% #we need to get rid of salinity -> TempComp
  filter(chamber == 1) %>% 
  select(-(Salinity:TempComp)) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(O_perecent_airsat:Temperature_compensation_C, chamber:seconds)


Chamber2_8 <- FR_2 %>% 
  filter(chamber > 1) %>% 
  select(-(Salinity_N:Temperature_compensation_C)) %>% 
  relocate(Salinity:TempComp, .before = PowerX4_EAC7_voltage_V_flush) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Oairsat:TempComp, chamber:seconds)

colnames(Chamber1) <- colnames(Chamber2_8) #renaming all of the columns in the Chamber1 dataframe with the columns names from Chamber2_8
FieldResp <- rbind(Chamber1, Chamber2_8) #Combining the two dataframes as FieldResp
write.csv(FieldResp, "FieldResp_2026-02-18.csv")

FieldResp %>% 
  ggplot(aes(x = seconds, y = Oairsat, color = chamber)) +
  geom_point()
```

### RespR Package

```{r}
# Convert your oxygen data to a respR object
inspect(FieldResp)
```

### Calculate the Slope

```{r}
#Filtering for only chamber 1
Ch1 <- FieldResp %>% 
  filter(chamber == 1)

Ch1_mmr <- calc_rate(Ch1,
                     from = 359, # three minutes 'wait' after start
                     to = 27838,   # five minutes later = 'measure' phase
                     by = "row") |>
  summary()

Ch1_conv <- convert_rate(Ch1,
                         oxy.unit = "mg/L",       # oxygen units of the original raw data
                         time.unit = "secs",      # time units of the original raw data
                         output.unit = "mg/h/g",  # desired output unit
                         volume = 0.74392,        # effective volume of the respirometer in L
                         mass = 0.02058)          # mass of the specimen in kg

summary(Ch1_conv)
```

### Playing with Urchin Galapagos Data

```{r}
# define the Sharpe-Schoolfield equation
schoolfield_high <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) #units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)
}
# fit over each set of groupings

#droplevels(mydata$Organism.ID)

mydata$Location<-as.character(mydata$Location)
mydata$Organism.ID<-as.character(mydata$Organism.ID)

fits <- mydata %>%
  group_by(Organism.ID, Location) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ schoolfield_high(lnc, E, Eh, Th, temp = K, Tc = 26),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))

```
