---
title: "Final Project DS4EEB"
author: "Hannah Hayes"
format: html
editor: visual
---

### Libraries & Packages

```{r}
#install.packages("usethis")
library("usethis")
#install.packages("here")
library("here")
#install.packages("dplyr")
library(tidyverse) #this will help us clean up the data 
library(readr)
#install.packages("respR")
library(respR)
#install.packages("devtools")
library(devtools)
library(lubridate) #To turn time HH:MM:SS AM/PM into seconds passed
#install.packages("rTPC")
library(rTPC)
library(dplyr)
library(ggplot2)
```

### Working Directory

```{r}
#file.choose()
setwd("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry") 
```

### Read in the Data

The way that LoLigo has give us the data is a separate CSV for each chamber, and they don't have a column for which chamber it is. So I'm going to read in all the data and add a column for the chamber number.

```{r}
#Field Resp Chamber 1 
file.choose()
FRCh1 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber1_218202610216_UTF.csv")
FRCh1 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh1 <- FRCh1 %>%
  select(-Date) %>% 
 mutate(chamber = 1, Date = "2026-02-18")

#Field Resp Chamber 2 
file.choose()
FRCh2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Rcode/CSUMB-UCSC-Urchin-Respirometry/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber2_218202610216_UTF.csv")
FRCh2 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh2 <- FRCh2 %>%
  select(-Date) %>% 
 mutate(chamber = 2, Date = "2026-02-18")

#Field Resp Chamber 3
FRCh3 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber3_218202610216_UTF.csv")
FRCh3 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh3 <- FRCh3 %>%
 mutate(chamber = 3, Date = "2026-02-18")

#Field Resp Chamber 4
FRCh4 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber4_218202610216_UTF.csv")
FRCh4 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh4 <- FRCh4 %>%
 mutate(chamber = 4, Date = "2026-02-18")

#Field Resp Chamber 5
FRCh5 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber5_218202610216_UTF.csv")
FRCh5 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh5 <- FRCh5 %>%
 mutate(chamber = 5, Date = "2026-02-18")

#Field Resp Chamber 6
FRCh6 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber6_218202610216_UTF.csv")
FRCh6 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh6 <- FRCh6 %>%
 mutate(chamber = 6, Date = "2026-02-18")

#Field Resp Chamber 7
FRCh7 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber7_218202610216_UTF.csv")
FRCh7 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh7 <- FRCh7 %>%
 mutate(chamber = 7, Date = "2026-02-18")

#Field Resp Chamber 8
FRCh8 <- read.csv("~/Desktop/JJG x HH Collab/Rcode/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber8_218202610216_UTF.csv")
FRCh8 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh8 <- FRCh8 %>%
 mutate(chamber = 8, Date = "2026-02-18")

```

### Combine all of the CSVs

```{r}
FR <- bind_rows(FRCh1, FRCh2, FRCh3, FRCh4, FRCh5, FRCh6, FRCh7, FRCh8) #Combine all of the data to one CSV
FR #Check that it worked

unique(FR$chamber) #Check that all the chambers made it into the same CSV
```

### Convert AM/PM into 24hr

```{r}
# Call upon Time column 
FR$Time

# Convert to a time object with a temporary date, recognizing the AM/PM format
# strptime() automatically converts to 24-hour format internally
time_objects <- strptime(FR$Time, format = "%I:%M:%S %p")

# Format the time objects to display only the time in 24-hour format (military time)
FR$Time <- format(time_objects, format = "%H:%M:%S")

# Print the result
FR$Time
```

### Convert Time into Seconds

```{r}
FR_sec <- FR %>%
  mutate(
    time_parsed = hms(Time),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Select for variables

```{r}
#O2%, Chamber, Date, seconds, temperature
# Select specific columns
FR_resp <- select(FR_sec, chamber, Date, )
# Or use the pipe operator (%>%) which is common in Tidyverse
FR_resp <- FR_sec %>% 
  select(column_name_1, column_name_2)
FR_resp$colname
```

#### Change Time into seconds

```{r}
#| eval: false
#| include: false
FR_sec <- FR %>%
  mutate(
    time_parsed = parse_date_time(Time, "I:M:S"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )

# Source - https://stackoverflow.com/a/31549819
# Posted by inhuretnakht, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-06, License - CC BY-SA 4.0

library(lubridate)
FR_sec <- FR %>% 
period_to_seconds(hms("10:02:17"))

FR_sec <- FR %>%
  filter(is.na(parse_date_time(Time, orders = "I:M:S p"))) %>%
  head()

FR_Time <- trimws(FR, which = Time)

FR_sec <- FR_Time %>%
  mutate(
    time_parsed = parse_date_time(Time, orders = "I:M:S p"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Duplicate NA Columns

```{r}
FR_sec
write.csv(FR_sec, "FR_sec.csv")
file.choose()
library(readr)
FR_2 <- read_csv("Data/FieldRep_021826-2/UTF/FR.csv")
#FR_2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FR.csv")

Chamber1 <- FR_2 %>% #we need to get rid of salinity -> TempComp
  filter(chamber == 1) %>% 
  select(-(Salinity:TempComp)) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Time, Intermittent_phase, O_perecent_airsat:Temperature_compensation_C, chamber:seconds)


Chamber2_8 <- FR_2 %>% 
  filter(chamber > 1) %>% 
  select(-(Salinity_N:Temperature_compensation_C)) %>% 
  relocate(Salinity:TempComp, .before = PowerX4_EAC7_voltage_V_flush) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Time, Intermittent_phase, Oairsat:TempComp, chamber:seconds)

colnames(Chamber1) <- colnames(Chamber2_8) #renaming all of the columns in the Chamber1 dataframe with the columns names from Chamber2_8
FieldResp <- rbind(Chamber1, Chamber2_8) #Combining the two dataframes as FieldResp
write.csv(FieldResp, "FieldResp_2026-02-18.csv")

FieldResp <- read.csv("FieldResp_2026-02-18.csv")

FieldResp %>% 
  ggplot(aes(x = seconds, y = Oairsat, color = chamber)) +
  geom_point()


Field_Resp_select <- FieldResp %>% 
  mutate(Tub = ifelse(chamber <= 4, "A", "B")) %>% 
  select(Time, Intermittent_phase, Oairsat,Temp, chamber, Date, seconds, Tub) 



#Group chamber 1-3 & 5-7 separately 


```

# STARTING OVER

### Data Wrangling

```{r}
FR1_8 <- read.csv("FieldResp_2026-02-18.csv") #Read in the data
#lets get all the column names 
colnames(FR1_8)

FR1_8 <- FR1_8 %>% #select for O ar saturation, chamber, temperautre, seconds
  select(Oairsat, Temp,chamber, seconds) %>% 
  #add in mass for each urchin
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/1000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/1000
    )
  ) %>% #Then we need to add in chamber volume 
   mutate( 
    ChV_ml = case_when(
      chamber == 1 ~ 743.92,
      chamber == 2 ~ 751.42,
      chamber == 3 ~ 743.96,
      chamber == 4 ~ 763.50,
      chamber == 5 ~ 751.16, 
      chamber == 6 ~ 745.90, 
      chamber == 7 ~ 747.43, 
      chamber == 8 ~ 763.50
    )
  ) %>% #add in a blank 1 & blank 2 column 
  mutate(
    blank = case_when(
      chamber <= 4 ~ 1,
      chamber >= 5 ~ 2
    )
  )
      
head(FR1_8) #check out the data to see if selected correctly 
tibble(FR1_8)
```

### 3. Background Noise

When doing respirometry, you typically run a blank chamber that does not have an organism in it to account for any background noise (respiration, photosynthesis) that may be affecting the O2 measurements that cannot be attributed to your study organism. We must calculate the oxygen consumption rates of the blanks and minus them from the actual rates from the other chambers.

### Oxygen Consumption

```{r}
FRplot <- FRresp %>% 
  ggplot(aes(
    x= Temp, y = oxygen
  )) +
  geom_point() 
FRplot

FRplot2 <- FRresp_grouped %>% 
  ggplot(aes(
    x= Temp, y = oxygen_rate
  )) +
  geom_point() 
FRplot2
```

### Let's add the calculations and adjustments to the FRresp

```{r}
#we need to add a temperature bin to the data fram 
FRresp <- FRresp %>% 
  mutate(
    temp_bin = round(Temp)
  )
  

#the we need to group everything into temo_bin and avergae it for each temperature step

#going to group by blank > chamber > time 

FRresp_grouped <- FRresp %>%
  arrange(blank, chamber, time) %>%   # sort first by blank,chamber, then time
  group_by(blank, chamber) %>%        # calculate rates within each replicate
  mutate(umol.sec = c(NA, diff(oxygen) / diff(time))) #%>% 
 # ungroup() %>%
 # mutate(blank.rate = ifelse(chamber == 4, umol.sec, NA))
```

### Starting Over Again

```{r}

# change in oxygen
delta_O2 <- O2_start - O2_end

# metabolic rate
MR <- (Oairsat2 / timepassed) * ChV_ml

freaky <- FR1_8 %>% #select for O ar saturation, chamber, temperautre, seconds
  select(Oairsat, Temp,chamber, time_parsed, seconds) %>% 
  #add in mass for each urchin
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/10000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/10000
    )
  ) %>% #Then we need to add in chamber volume 
   mutate( 
    ChV_ml = case_when(
      chamber == 1 ~ 743.92,
      chamber == 2 ~ 751.42,
      chamber == 3 ~ 743.96,
      chamber == 4 ~ 763.50,
      chamber == 5 ~ 751.16, 
      chamber == 6 ~ 745.90, 
      chamber == 7 ~ 747.43, 
      chamber == 8 ~ 763.50
    )
  ) %>% #add in a blank 1 & blank 2 column 
  mutate(
    blank = case_when(
      chamber <= 4 ~ 1,
      chamber >= 5 ~ 2
    )
  ) %>% 
  mutate(timepassed = 1) %>% 
  #need to add a column for change in oxygen o2 end 
  mutate(Oairsat2 = c(NA, diff(Oairsat) / diff(seconds))) 
   
      
head(freaky) #check out the data to see if selected correctly 
tibble(freaky)

# change in oxygen
FKEY_oxy_chg  <- freaky %>%
  mutate(
    delta_O2 = Oairsat2,
    metabolic_rate = ((delta_O2 / masskg * timepassed) 
  )

```

### calculate background rate

```{r}
#this is where we need to adjust for the rate 
#the we need to group everything into temo_bin and avergae it for each temperature step
#going to group by blank > chamber > time 

#Get the blank rates 
blank1_rate <- FKEY_oxy_chg %>%
  filter(chamber == 4) %>%
  pull(Oairsat2) %>%
  first()

blank2_rate <- FKEY_oxy_chg %>%
  filter(chamber == 8) %>%
  pull(Oairsat2) %>%
  first()

#Create the blank_rate column 
FKEY_adj <- FKEY_oxy_chg %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ blank1_rate,
      chamber %in% 5:7 ~ blank2_rate,
      TRUE ~ NA_real_
    )
  )

#Minus the MR of the background 
FKEY_adj <- FKEY_adj %>% 
  mutate(O2.sec.corr = Oairsat2 - blank_rate) %>% 
  filter(!chamber == "4" | !chamber == "8") 
################################################################################
#lets try grouping it here 
FKEY_adj_group <- FKEY_adj %>% 
   filter(is.finite(O2.sec.corr)) %>% 
  mutate(Temp = round(Temp, 1)) %>%
  group_by(Temp, chamber, masskg) %>%
  summarise(O2.sec.corr.mean = mean(O2.sec.corr)) 

#find the mml cm2 hr
FKEY_adj_group_hr <- FKEY_adj_group %>% 
  mutate(umol.cm2.hr = (O2.sec.corr.mean*3600)/masskg
  ) %>% 
  filter(umol.cm2.hr > 0)

#log the rates
FKEY_adj_group_log <- FKEY_adj_group_hr %>% 
mutate(rate.in = log(umol.cm2.hr+0.1))

FKEY_adj_group_clean <- FKEY_adj_group_log %>% 
  filter(is.finite(rate.in))
#########################################################################################

#find the mml cm2 hr
FKEY_adj_hr <- FKEY_adj %>% 
  mutate(umol.cm2.hr = (O2.sec.corr*3600)/masskg
  )

#remove all na rows
FKEY_hr_na<- FKEY_adj_hr %>% 
  filter(is.finite(blank_rate))

#group it by temp 
FKEY_grouped <- FKEY_hr_na %>%
  group_by(Temp) %>%
  summarise(metabolic_rate = mean(umol.cm2.hr), .groups = "drop") %>% 
  filter(is.finite(metabolic_rate))

#log the rates
FKEY_log <- FKEY_hr_na %>% 
mutate(rate.in = log(umol.cm2.hr+0.1))
#filter out NAs
FKEY_log_clean <- FKEY_log %>% 
  filter(is.finite(rate.in))
```

### Now we can calculate MR

```{r}
#recalculate MR
FKEY_MR  <- FKEY_adj %>%
  mutate(
    delta_O2 = Oairsat2,
    metabolic_rate = (O2.sec.corr / masskg * timepassed) 
  )

#now lets group it by temperature 
FKEY_avg <- FKEY_MR %>%
  group_by(Temp) %>%
  summarise(metabolic_rate = mean(metabolic_rate), .groups = "drop")


#account for mass
#FKEY_mass <- FKEY_oxy_chg %>%
 # mutate(
  #  mass_specific_MR = metabolic_rate / masskg
  #)

#need the slope of oxygen decline 
#model <- lm(Oairsat ~ seconds, data = FKEY_mass)

#slope <- coef(model)[2]

#add MR as a column
#FKEY_MR <- FKEY_mass %>% 
 # mutate(metabolic_rate = slope * ChV_ml
 # )
#head(FKEY_MR)
```

#### Quick Visualizations

```{r}
MRplot <- FKEY_grouped %>% 
  ggplot(aes(x= Temp, y = metabolic_rate)) +
  geom_point()
MRplot

plot(FKEY_adj_group_log $Temp,
     FKEY_adj_group_log $rate.in,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")

plot(FKEY_adj_group_hr$Temp,
     log(FKEY_adj_group_hr$umol.cm2.hr),
     pch = 16,
     xlab = "Temperature",
     ylab = "log(Metabolic Rate)")
```

### 

### Blanks Data

```{r eva; = FALSE}
#make a new data frame with only the blanks chamber
blank_data <- FKEY_MR %>%
  filter(chamber == 4 | chamber == 8)

#adjust for background rate 
blank_model <- lm(Oairsat ~ seconds, data = blank_data)
blank_slope <- coef(blank_model)[2]



```

### Model

```{r}
#remove Nas
FKEY_MR <- FKEY_MR %>%
  filter(is.finite(metabolic_rate), is.finite(Temp))

#fill in missing values
start_vals["el"] <- 1
#start_vals["eh"] <- 1

#recalculate starting values
start_vals <- get_start_vals(
  FKEY_MR$Temp,
  FKEY_MR$metabolic_rate,
  model_name = "sharpeschoolfull_1981"
)

#check vals
start_vals
is.finite(start_vals)

#check for negative or zero rates
summary(FKEY_MR$metabolic_rate)

#fit the model 
mod <- nls_multstart(
  metabolic_rate ~ sharpeschoolfull_1981(
    temp = Temp,
    r_tref,
    e,
    el,
    tl,
    eh,
    th,
    tref = 20
  ),
  data = FKEY_MR,
  iter = c(3,3,3,3,3,3),
  start_lower = start_vals - 10,
  start_upper = start_vals + 10,
  lower = get_lower_lims(FKEY_MR$Temp, FKEY_MR$metabolic_rate,
                         model_name = "sharpeschoolfull_1981"),
  upper = get_upper_lims(FKEY_MR$Temp, FKEY_MR$metabolic_rate,
                         model_name = "sharpeschoolfull_1981"),
  supp_errors = "Y"
)

#Check the fit 
summary(mod)
#generate predictions for plotting 
preds <- data.frame(
  Temp = seq(min(tpc_data$Temp), max(tpc_data$Temp), length.out = 100)
)
preds <- broom::augment(mod, newdata = preds)

#plot TPC 
ggplot() +
  geom_point(data = tpc_data, aes(Temp, rate_mass)) +
  geom_line(data = preds, aes(Temp, .fitted), color = "blue") +
  theme_bw()

summary(tpc_data$rate_mass)

summary(tpc_data)
summary(mod)
```

### I had a dream, lets calculate oxygen consumed for every

```{r}
#we're working with the field resp select data frame because I need the loop column 
Field_Resp_select

 # we need these columns timepassed = 1, Oairsar2, tempK, metabolic rate, blank rate, o2seccorrected, o2 mg hour
Field_Reskp_grouped <- Field_Resp_select %>%
  arrange(Tub, chamber, Time) %>%   # sort first by blank,chamber, then time
  group_by(Tub, chamber) %>%        # calculate rates within each replicate
  mutate(umol.sec = c(NA, diff(Oairsat) / diff(seconds))) #%>% 
 # ungroup() %>%
 # mutate(blank.rate = ifelse(chamber == 4, umol.sec, NA)) 

#Now we need the blank rate to minus 
#Get the blank rates 
blank1_rate <- Field_Reskp_grouped %>%
  filter(chamber == 4) %>%
  pull(umol.sec)
  

blank2_rate <- Field_Reskp_grouped %>%
  filter(chamber == 8) %>%
  pull(umol.sec)

#Create the blank_rate column 
Field_Reskp_adj <- Field_Reskp_grouped %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ blank1_rate,
      chamber %in% 5:7 ~ blank2_rate,
      TRUE ~ NA_real_
    )
  )%>% 
  filter(!chamber == "4" | !chamber == "8") 

#seperate m from numbers in the intermittent phase column
Field_Reskp_adj <- Field_Reskp_adj %>%
  mutate(Intermittent_phase = as.numeric(gsub("M", "", Intermittent_phase)))

#Lets add in a temp bin
Field_Reskp_tbin <- Field_Reskp_adj %>% 
mutate(temp_bin = round(Temp))

###########################################################################
#playing to see if it's better to group things by the loop #
###########################################################################

#lets group it by measuremnet phase and take the average of that 
Field_Resp_loop <- Field_Reskp_adj %>%
  group_by(Intermittent_phase) %>% # calculate rates within each replicate
  summarise(mean_rate = mean(O2.sec.corr, na.rm = TRUE))

#Let's plot to see what it looks like   
  plot(Field_Resp_loop$Intermittent_phase,
     Field_Resp_loop$mean_rate,
     pch = 16,
     xlab = "Loop #",
     ylab = "Average O2 Consumption",
     main = "TPC diagnostic plot")
#it appears loop# is not the right way to group it. Makes our sample size too small
```

### Lets try the same grouping, but by temperature instead of phase

```{r}
#Lets make a column that rounds the temp up or down

#lets group it by measuremnet phase and take the average of that 
Field_Reskp_tempavg <- Field_Reskp_adj %>%
  mutate(temp_bin = round(Temp)) %>% 
  group_by(temp_bin) %>% # calculate rates within each replicate
  summarise(mean_rate = mean(O2.sec.corr, na.rm = TRUE))

 #Let's plot to see what it looks like   
  plot(Field_Reskp_tempavg$temp_bin,
     Field_Reskp_tempavg$mean_rate,
     pch = 16,
     xlab = "Temp",
     ylab = "Average O2 Consumption",
     main = "TPC diagnostic plot")
  
#looks nice! Let's make a datafram with everything including temp_bin
Field_Reskp_temp <- Field_Reskp_adj %>%
  mutate(temp_bin = round(Temp))
```

### Now that the data looks nice let's factor in mass

```{r}
#Add a mass column and chamber volume
Field_Reskp_mass_chV <- Field_Reskp_temp %>% 
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/1000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/1000
    )
  ) %>% #Then we need to add in chamber volume 
   mutate( 
    ChV_L = case_when(
      chamber == 1 ~ 743.92/1000,
      chamber == 2 ~ 751.42/1000,
      chamber == 3 ~ 743.96/1000,
      chamber == 4 ~ 763.50/1000,
      chamber == 5 ~ 751.16/1000, 
      chamber == 6 ~ 745.90/1000, 
      chamber == 7 ~ 747.43/1000, 
      chamber == 8 ~ 763.50/1000
    )) %>% #filter out NAs
    filter(!is.na(O2.sec.corr))  

arrange()

#I think I need to adjust for volume 
#Field_Reskp_mass_vol <- Field_Reskp_mass_chV %>% 
 # mutate(umol.sec = umol.sec*ChV_L)

#find the mml cm2 hr


Field_Reskp_mmlcm2hr <- Field_Reskp_mass_chV %>% 
  mutate(umol.cm2.hr = (O2.sec.corr)/masskg
  ) %>% 
  filter(umol.cm2.hr > 0)
```

### aggregate function

```{r}
#We're going to group it differently this time
aggy <- aggregate(umol.sec ~ Tub*temp_bin*chamber, data=Field_Reskp_tbin, mean)
# pull out only the blanks
#photo.blnk<-photo.blnk[photo.blnk$Species=='BK',]
aggy <-aggy[aggy$Tub==4,]
#joinit back to the tbin data frame
aggy$Tub<-NULL
colnames(aggy)[4]<-'blank.rate' # rename the blank rate 
# join the blank data with the rest of the data
Photo.R<-left_join(Photo.R, photo.blnk)

#minus the blanks 
blank1_rate <- aggy %>%
  slice(-148) %>%
  filter(chamber == 4) %>%
  pull(umol.sec)
  

blank2_rate <- aggy %>%
  filter(chamber == 8) %>%
  pull(umol.sec)

#Create the blank_rate column 
aggy_adj <- aggy %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ first(blank1_rate),
      chamber %in% 5:7 ~ first(blank2_rate),
      TRUE ~ NA_real_
    )
  )#%>% 
  #filter(!chamber == "4" | !chamber == "8") 

#Minus the blank rate and the umoll to get the adjusted value
aggy_corr <- aggy_adj %>% 
  mutate(O2.sec.corr = umol.sec - blank_rate) 

#Correct for the hour, but we need the mass
aggy_hh <- aggy_corr %>% 
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/1000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/1000
    )
  ) %>% 
  mutate(umol_cm2_hr = (O2.sec.corr * 3600) / masskg)

#remove NAs
aggy_na <- aggy_hh %>% 
    filter(!is.na(umol_cm2_hr))
#make it positive
  aggy_pos <- aggy_na %>% 
    mutate(umol_cm2_hr = -(umol_cm2_hr)) %>% 
    filter(umol_cm2_hr > 0)
  
#log it for get the rate
   aggy_log <- aggy_pos %>% 
    mutate(rate = log(umol_cm2_hr + 0.01))
   
#we fianlly got some good data!
write.csv(aggy_log, "TPCdata.csv")
```

### Lets check the visuals

```{r}
#logged Rate vs temp bin 
plot(aggy_log $temp_bin,
     aggy_log $rate,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")

plot(aggy_log $temp_bin,
     aggy_log $umol_cm2_hr,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")
```

## Models

### Read in the data

```{r}
#read in the data
TPCdata <- read.csv("TPCdata.csv")
```

### BRMS

```{r}
fit_simple <- 
  brm(data = TPCdata, # Give the model the data
      # Choose a gaussian (normal) distribution
      family = gaussian,
      # Specify the model here. 
      rate ~ 1 + temp_bin,
      # Here's where you specify parameters for executing the Markov chains
      # We're using similar to the defaults, except we set cores to 4 so the analysis runs faster than the default of 1
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      # Setting the "seed" determines which random numbers will get sampled.
      # In this case, it makes the randomness of the Markov chain runs reproducible 
      # (so that both of us get the exact same results when running the model)
      seed = 4,
      # Save the fitted model object as output - helpful for reloading in the output later
      file = "output/fit_simple")

summary(fit_simple)
plot(fit_simple)
```

### I noticed in the histogram the data looks a bit skewed so I would like ot try using a gamma family instead of a Guassain

```{r}
fit_gamma <- 
  brm(data = TPCdata, # Give the model the data
      # Choose a gaussian (normal) distribution
      family = Gamma(),
      # Specify the model here. 
      rate ~ 1 + temp_bin,
      # Here's where you specify parameters for executing the Markov chains
      # We're using similar to the defaults, except we set cores to 4 so the analysis runs faster than the default of 1
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      # Setting the "seed" determines which random numbers will get sampled.
      # In this case, it makes the randomness of the Markov chain runs reproducible 
      # (so that both of us get the exact same results when running the model)
      seed = 4,
      # Save the fitted model object as output - helpful for reloading in the output later
      file = "output/fit_gamma")

summary(fit_gamma)
plot(fit_gamma)
```

### Let's visualize the data

```{r}
model <- TPCdata %>% 
  ggplot(aes(x = temp_bin, y = rate)) +
  geom_point() 
```

### Is it normal data? 

```{r}
hist <- TPCdata %>% 
  ggplot(aes(x = rate)) +
  geom_histogram()
hist

hist2 <- TPCdata %>% 
  ggplot(aes(x = umol_cm2_hr)) +
  geom_histogram()
hist2
```
