---
title: "Final Project DS4EEB"
author: "Hannah Hayes"
format: html
editor: visual
---

### Libraries & Packages

```{r}
#install.packages("usethis")
library("usethis")
#install.packages("here")
library("here")
#install.packages("dplyr")
library(tidyverse) #this will help us clean up the data 
library(readr)
#install.packages("respR")
library(respR)
#install.packages("devtools")
library(devtools)
library(lubridate) #To turn time HH:MM:SS AM/PM into seconds passed
```

### Working Directory

```{r}
#file.choose()
setwd("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF") 
```

### Read in the Data

The way that LoLigo has give us the data is a separate CSV for each chamber, and they don't have a column for which chamber it is. So I'm going to read in all the data and add a column for the chamber number.

```{r}
#Field Resp Chamber 1 
FRCh1 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber1_218202610216_UTF.csv")
FRCh1 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh1 <- FRCh1 %>%
  select(-Date) %>% 
 mutate(chamber = 1, Date = "2026-02-18")

#Field Resp Chamber 2 
FRCh2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber2_218202610216_UTF.csv")
FRCh2 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh2 <- FRCh2 %>%
  select(-Date) %>% 
 mutate(chamber = 2, Date = "2026-02-18")

#Field Resp Chamber 3
FRCh3 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber3_218202610216_UTF.csv")
FRCh3 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh3 <- FRCh3 %>%
 mutate(chamber = 3, Date = "2026-02-18")

#Field Resp Chamber 4
FRCh4 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber4_218202610216_UTF.csv")
FRCh4 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh4 <- FRCh4 %>%
 mutate(chamber = 4, Date = "2026-02-18")

#Field Resp Chamber 5
FRCh5 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber5_218202610216_UTF.csv")
FRCh5 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh5 <- FRCh5 %>%
 mutate(chamber = 5, Date = "2026-02-18")

#Field Resp Chamber 6
FRCh6 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber6_218202610216_UTF.csv")
FRCh6 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh6 <- FRCh6 %>%
 mutate(chamber = 6, Date = "2026-02-18")

#Field Resp Chamber 7
FRCh7 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber7_218202610216_UTF.csv")
FRCh7 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh7 <- FRCh7 %>%
 mutate(chamber = 7, Date = "2026-02-18")

#Field Resp Chamber 8
FRCh8 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FieldRep_021826_Restingchamber8_218202610216_UTF.csv")
FRCh8 #Check if it read in correctly

  ##We're going to overwrite FRCh1 so it includes the chamber column
FRCh8 <- FRCh8 %>%
 mutate(chamber = 8, Date = "2026-02-18")

```

### Combine all of the CSVs

```{r}
FR <- bind_rows(FRCh1, FRCh2, FRCh3, FRCh4, FRCh5, FRCh6, FRCh7, FRCh8) #Combine all of the data to one CSV
FR #Check that it worked

unique(FR$chamber) #Check that all the chambers made it into the same CSV
```

### For Loop

```{r}

```

### Convert AM/PM into 24hr

```{r}
# Call upon Time column 
FR$Time

# Convert to a time object with a temporary date, recognizing the AM/PM format
# strptime() automatically converts to 24-hour format internally
time_objects <- strptime(FR$Time, format = "%I:%M:%S %p")

# Format the time objects to display only the time in 24-hour format (military time)
FR$Time <- format(time_objects, format = "%H:%M:%S")

# Print the result
FR$Time
```

### Convert Time into Seconds

```{r}
FR_sec <- FR %>%
  mutate(
    time_parsed = hms(Time),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Select for variables

```{r}
#O2%, Chamber, Date, seconds, temperature
# Select specific columns
FR_resp <- select(FR_sec, chamber, Date, )
# Or use the pipe operator (%>%) which is common in Tidyverse
FR_resp <- FR_sec %>% 
  select(column_name_1, column_name_2)
FR_resp$colname
```

### Trials

#### Trial #1

okay so we're having way more trouble loading in the data. I think we may have got it now so let's keep going. So through a lot of trial and error and googling and using chatgpt. Turns out the orgiianl file was a UTF 16, but r uses a UTF -8 files. So using TextEdit I was able to change the type of file then i reopened it in ecel and turned it into a CSBV now it works!!!!!!

#### Trial #2

Don't forget to check that everything is there. I didn't notice till after all of these steps that the seconds were missing from the time column so I have to go back and fix all of the csv files again!!!!

```         
```

### Snips

#### 1st attempt at bindrows

This was to bind all of the CSVs from my local file together, but then I forgot none of them actually have the chamber column

`{r| eval: false} #FR <- list.files(   #path = "/Users/hannahhayes858/Desktop/JJG x HH Collab/CSUMB-UCSC-Urchin-Respirometry/FieldResp_021826/UT8s/TestingforUTF",   #full.names = TRUE #) %>%   # lapply(read_csv) %>%   # bind_rows() #FR #write.csv(FR, "fieldrespch1_8.csv")`

#### Separate Date and Time

\`\`\`{r\| eval: false} #First we need to seperate the date and the time

df_char \<- data.frame( datetime_string = c("2021-05-25 12:34:25", "2019-01-14 04:10:30"))

df_separated \<- FR %\>% \# separate(Datetime, into = c("date", "time"), sep = " ") df_separated \`\`\`

#### Sardine Data Set Example

\`\`\`{r\| eval: false} #\| eval: false #\| include: false \# Example using built-in data 'sardine.rd' data(sardine.rd) \# Inspect the data (time in column 1, oxygen in column 2) sardine_insp \<- inspect(sardine.rd, time = 1, oxygen = 2)

Calculate rate from a specific region (manual)

manual_rate \<- calc_rate(sardine_insp, from = 1000, to = 3000)

Automatically detect the "best-fit" linear rate

auto_rate_obj \<- auto_rate(sardine_insp, method = "linear")

Calculate background rate from background data (example data not shown)

bg_rate \<- calc_rate.bg(background_data)

Adjust the specimen rate

adjusted_rate \<- adjust_rate(auto_rate_obj, by = bg_rate)

Convert to mass-specific units (requires specimen mass and chamber volume)

final_rate \<- convert_rate(adjusted_rate, volume = 1.2, \# volume of respirometer in L mass = 0.5, \# mass of specimen in kg \# input/output units will depend on your data O2.unit = "mg/L", time.unit = "hours", output.unit = "mg/h/kg")

```{r}
#devtools::install_github("januarharianto/respR")


#load the data and combine all csv files
# Source - https://stackoverflow.com/a/30242859
# Posted by Maiasaura, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-04, License - CC BY-SA 4.0
```

#### Change Time into seconds

```{r}
#| eval: false
#| include: false
FR_sec <- FR %>%
  mutate(
    time_parsed = parse_date_time(Time, "I:M:S"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )

# Source - https://stackoverflow.com/a/31549819
# Posted by inhuretnakht, modified by community. See post 'Timeline' for change history
# Retrieved 2026-03-06, License - CC BY-SA 4.0

library(lubridate)
FR_sec <- FR %>% 
period_to_seconds(hms("10:02:17"))

FR_sec <- FR %>%
  filter(is.na(parse_date_time(Time, orders = "I:M:S p"))) %>%
  head()

FR_Time <- trimws(FR, which = Time)

FR_sec <- FR_Time %>%
  mutate(
    time_parsed = parse_date_time(Time, orders = "I:M:S p"),
    seconds = as.numeric(time_parsed - first(time_parsed))
  )
```

### Duplicate NA Columns

```{r}
FR_sec
write.csv(FR_sec, "FR_sec.csv")
file.choose()
FR_2 <- read.csv("/Users/hannahhayes858/Desktop/JJG x HH Collab/Data/FieldRep_021826-2/UTF/FR.csv")

Chamber1 <- FR_2 %>% #we need to get rid of salinity -> TempComp
  filter(chamber == 1) %>% 
  select(-(Salinity:TempComp)) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(O_perecent_airsat:Temperature_compensation_C, chamber:seconds)


Chamber2_8 <- FR_2 %>% 
  filter(chamber > 1) %>% 
  select(-(Salinity_N:Temperature_compensation_C)) %>% 
  relocate(Salinity:TempComp, .before = PowerX4_EAC7_voltage_V_flush) %>% 
  filter(Intermittent_mode == "RITM") %>% 
  select(Oairsat:TempComp, chamber:seconds)

colnames(Chamber1) <- colnames(Chamber2_8) #renaming all of the columns in the Chamber1 dataframe with the columns names from Chamber2_8
FieldResp <- rbind(Chamber1, Chamber2_8) #Combining the two dataframes as FieldResp
write.csv(FieldResp, "FieldResp_2026-02-18.csv")

FieldResp <- read.csv("FieldResp_2026-02-18.csv")

FieldResp %>% 
  ggplot(aes(x = seconds, y = Oairsat, color = chamber)) +
  geom_point()



Field_Resp_select <- FieldResp %>% 
  mutate(Tub = ifelse(chamber <= 4, "A", "B")) %>% 
  select(Oairsat,Temp, chamber, Date, seconds, Tub) 



#Group chamber 1-3 & 5-7 separately 


```

### RespR Package

```{r}
# Convert your oxygen data to a respR object
inspect(FieldResp)
```

### Calculate the Slope

```{r}
#Filtering for only chamber 1
Ch1 <- FieldResp %>% 
  filter(chamber == 1)

Ch1_mmr <- calc_rate(Ch1,
                     from = 359, # three minutes 'wait' after start
                     to = 27838,   # five minutes later = 'measure' phase
                     by = "row") |>
  summary()

Ch1_conv <- convert_rate(Ch1,
                         oxy.unit = "mg/L",       # oxygen units of the original raw data
                         time.unit = "secs",      # time units of the original raw data
                         output.unit = "mg/h/g",  # desired output unit
                         volume = 0.74392,        # effective volume of the respirometer in L
                         mass = 0.02058)          # mass of the specimen in kg

summary(Ch1_conv)
```

### Playing with Urchin Galapagos Data

```{r}
# define the Sharpe-Schoolfield equation
schoolfield_high <- function(lnc, E, Eh, Th, temp, Tc) {
  Tc <- 273.15 + Tc
  k <- 8.62e-5
  boltzmann.term <- lnc + log(exp(E/k*(1/Tc - 1/temp))) #units are eV/K, electrovolts/Kelvin
  inactivation.term <- log(1/(1 + exp(Eh/k*(1/Th - 1/temp))))
  return(boltzmann.term + inactivation.term)
}
# fit over each set of groupings

#droplevels(mydata$Organism.ID)

mydata$Location<-as.character(mydata$Location)
mydata$Organism.ID<-as.character(mydata$Organism.ID)

fits <- mydata %>%
  group_by(Organism.ID, Location) %>%
  nest() %>%
  mutate(fit = purrr::map(data, ~ nls_multstart(log.rate ~ schoolfield_high(lnc, E, Eh, Th, temp = K, Tc = 26),
                                                data = .x,
                                                iter = 1000,
                                                start_lower = c(lnc = -10, E = 0.1, Eh = 0.2, Th = 285),
                                                start_upper = c(lnc = 10, E = 2, Eh = 5, Th = 330),
                                                supp_errors = 'Y',
                                                na.action = na.omit,
                                                lower = c(lnc = -10, E = 0, Eh = 0, Th = 0))))

```

# STARTING OVER

### Libraries & Packages

```{r}
#install.packages("usethis")
library("usethis") #to help use github
#install.packages("here")
library("here") #to help use github
#install.packages("dplyr")
library(tidyverse) #this will help us clean up the data 
library(readr)
#install.packages("respR")
library(respR) #to convert O2 into metabolic rate 
#install.packages("devtools")
library(devtools)
library(lubridate) #To turn time HH:MM:SS AM/PM into seconds passed
```

### Data Wrangling

#### !! STILL NEED TO ADD IN URCHIN ID !!

```{r}
FR1_8 <- read.csv("FieldResp_2026-02-18.csv") #Read in the data
#lets get all the column names 
colnames(FR1_8)

FR1_8 <- FR1_8 %>% #select for O ar saturation, chamber, temperautre, seconds
  select(Oairsat, Temp,chamber, seconds) %>% 
  #add in mass for each urchin
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/1000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/1000
    )
  ) %>% #Then we need to add in chamber volume 
   mutate( 
    ChV_ml = case_when(
      chamber == 1 ~ 743.92,
      chamber == 2 ~ 751.42,
      chamber == 3 ~ 743.96,
      chamber == 4 ~ 763.50,
      chamber == 5 ~ 751.16, 
      chamber == 6 ~ 745.90, 
      chamber == 7 ~ 747.43, 
      chamber == 8 ~ 763.50
    )
  ) %>% #add in a blank 1 & blank 2 column 
  mutate(
    blank = case_when(
      chamber <= 4 ~ 1,
      chamber >= 5 ~ 2
    )
  )
      
head(FR1_8) #check out the data to see if selected correctly 
tibble(FR1_8)
```

## RespR Package

### Data Wrangle

```{r}
#To use the respR package the data must have the "time" and "oxygen" in the first and second column respectively 

#Rename and move time (seconds) and oxygen columns (Oairsat)
FRresp <- FR1_8 %>% 
  relocate(seconds, Oairsat) %>% #Moves seconds and Oairsat to front
  rename(time = seconds, oxygen = Oairsat) #reames columns 

#we need to pivot wider the data so that each urchin has it's own oxygen column
FRwider <- FRresp %>%  
  pivot_wider(
     id_cols = time,
    names_from = chamber, 
    values_from = oxygen,
  ) %>% #Need to move blank from chamber 4 (Blank 1) before chamber 8 (Blank 2)
  relocate("4", .before = "8")
```

### 1. Inspect

```{r}
#First we must inspect the data for common issues 
urchin_int <- inspect(FRresp) #this is inspecting the entire data in the long foramt where there are only two columsn for time and oxygen 

#This will inspect the data for each indivudal chamber 
inspect(FRwider, time = 1, oxygen = 2:9)
```

### 2. Calc Rate

We then must calcultae the rate at which oxygen is being consumed. The respR package splits replicates up by the rows they are in. So we will use the longer format

```{r}
#Find from which row to which row is each replicate in
#using a for loop to keep reiterating the min and max of chamber rows 
for (i in 1:8) {
print(paste("Chamber", i))
  row_i <- which(FRresp$chamber == i)
  
  print(min(row_i))
  #print(max(row_i))

}
# Row 1 (1:17400)
# Row 2 (17401:34800)
# Row 3 (34801:52200)
# Row 4 (52201:69600)
# Row 5 (69601:87000)
# Row 6 (87001:104400)
# Row 7 (104401:121800)
# Row 8 (121801:139200)

#now we can calculate oxygen consumption rates 
urchin_rates <- calc_rate.int(FRresp,
              starts = c(1, 17401, 34801, 52201,69601, 87001, 104401, 121801)) 
summary(urchin_rates)
plot(urchin_rates)
```

### 3. Background Noise 

When doing respirometry, you typically run a blank chamber that does not have an organism in it to account for any background noise (respiration, photosynthesis) that may be affecting the O2 measurements that cannot be attributed to your study organism. We must calculate the oxygen consumption rates of the blanks and minus them from the actual rates from the other chambers.

```{r}
bg_insp <- inspect(FRwider, time = 1, oxygen = 8:9) # We inspect the data of only the blanks and we use this from the only the blanks 1 & 2 columns 
bg_rate <- calc_rate.bg(bg_insp) #then we calculate the background rates
print(bg_rate) #actual background rates were individually and then the mean 

#We then use it to adjust our urchin rates

urch_rates_adj <- adjust_rate(urchin_rates,
                              by = bg_rate) %>% 
  summary()
```

### Convert rates into units for metabolic rate 

```{r}
#We must take into consideration the mass of each urchin. so first we'll make a vector of the individual masses
mass_vec <- c(
  0.02058,  # urchin 1
  0.01308,  # urchin 2
  0.02054,  # urchin 3
  0.001,         # blank chamber 4
  0.01334,  # urchin 5
  0.0186,  # urchin 6
  0.01707,  # urchin 7
  0.001        # blank chamber 8
)

#do the same for volume
 volume <- c(
      743.92, #Ch 1
      751.42, #Ch 2
      743.96, # Ch 3
      763.50, # ch4
      751.16, # Ch 5
      745.90, # Ch 6
      747.43, #Ch 7
      763.50 #Ch 8 
 )

urch_rates_conv <- convert_rate(urch_rates_adj,
                                oxy.unit = "mg/l",
                                time.unit = "s",
                                output.unit = "mg/h/g",
                                mass = mass_vec,
                                volume = volume) 
class(urch_rates_conv)
nrow(urch_rates_conv)
```

### Prepping for Schoolfield Model 

```{r}
#we need to add a few columns to run model 

#umol.L.sec
#umol.sec
#Vol.L
#blank.rate	
#umol.sec.corr
#umol.cm2.hr
#Rate.ln


```

### Oxygen Consumption

```{r}
FRplot <- FRresp %>% 
  ggplot(aes(
    x= Temp, y = oxygen
  )) +
  geom_point() 
FRplot

FRplot2 <- FRresp_grouped %>% 
  ggplot(aes(
    x= Temp, y = oxygen_rate
  )) +
  geom_point() 
FRplot2
```

### Let's add the calculations and adjustments to the FRresp

```{r}
#we need to add a temperature bin to the data fram 
FRresp <- FRresp %>% 
  mutate(
    temp_bin = round(Temp)
  )
  

#the we need to group everything into temo_bin and avergae it for each temperature step

#going to group by blank > chamber > time 

FRresp_grouped <- FRresp %>%
  arrange(blank, chamber, time) %>%   # sort first by blank,chamber, then time
  group_by(blank, chamber) %>%        # calculate rates within each replicate
  mutate(umol.sec = c(NA, diff(oxygen) / diff(time))) #%>% 
 # ungroup() %>%
 # mutate(blank.rate = ifelse(chamber == 4, umol.sec, NA))

#Then add the blank calc rate 
df <- FRresp_grouped %>%
  group_by(blank, time) %>%
  mutate(blank_rate = if(any(chamber == 4)) umol.sec[chamber == 4] else NA_real_) %>%
  ungroup()


#Add in the blank rates column 
df2 <- FRresp_grouped %>%
  group_by(blank, time) %>%
  mutate(
    blank_rate = if(any(chamber == 4 & blank == 1)) umol.sec[chamber == 4 & blank == 1] else
                 if(any(chamber == 8 & blank == 2)) umol.sec[chamber == 8 & blank == 2] else NA_real_
  ) %>%
  filter_out(chamber == 4|8)
  ungroup()

# Row 1 (1:17400)
# Row 2 (17401:34800)
# Row 3 (34801:52200)
# Row 4 (52201:69600)
# Row 5 (69601:87000)
# Row 6 (87001:104400)
# Row 7 (104401:121800)
# Row 8 (121801:139200)

#mutate(blank.rate = umol.sec[chamber == 4][1]) #%>%  # pick chamber 4 as reference(

#Add in the adjusted calculated rates column 
df3 <- df2 %>% 
  mutate(umol.sec.corr = umol.sec - blank_rate) %>% 
  mutate(Oseckg = umol.sec.corr/masskg)  
#omg this has been a nightmare trying to filter out chambers 4 (blank 1) and 8 (blank 2)
write.csv(df3, "df3.csv")  
trythis <- read.csv("df3.csv")
trythis <- filter(trythis, !chamber == 4)
trythis2 <- filter(trythis, !chamber == 8)
write.csv(df2, "FRespcalculated.csv") #save it as a new csv
read.csv("FRespcalculated.csv")


#okay we also need the means of the rates THISISTHE NEW ONE
tpc_data2 <- trythis2 %>% 
  filter(!is.na(umol.sec.corr)) %>% 
  group_by(temp_bin) %>% 
  summarise(mean_rate = mean(Oseckg, na.rm = TRUE))


#remove NAs, mean for temperature step, group chamber and temp
tpc_data <- trythis3 %>%
  filter(!is.na(umol.sec.corr)) %>%
  group_by(temp_bin) %>%
  summarise(rate = mean(umol.sec.corr), .groups = "drop")

#tpcplot <- trythis3 %>% 
 # ggplot(aes(x = temp_bin, y = mean_rate)) +
 # geom_point()
#tpcplot

tpc_data2 <- tpc_data2 %>%
  mutate(rate_mass = rate / unique(trythis2$masskg))  
  


#install.packages("rTPC")
library(rTPC)
library(nls.multstart)
library(dplyr)
library(ggplot2)
#parameter estimates (THIS IS NOW ALREADY ACCOUNTING FoR MASS)
start_vals <- get_start_vals(tpc_data2$temp_bin, tpc_data2$mean_rate, 
                             model_name = "sharpeschoolfull_1981")
find("get_start_vals") #confirm it's loaded

#repeat for the rate accounted with mass 
get_start_vals(tpc_data$Temp, tpc_data$rate_mass,
               model_name = "sharpeschoolfull_1981")
start_vals["el"] <- 1 #fix those categorical values

#fit the model 
mod <- nls_multstart(
  mean_rate ~ sharpeschoolfull_1981(
    temp = temp_bin,
    r_tref,
    e,
    el,
    tl,
    eh,
    th,
    tref = 20
  ),
  data = tpc_data2,
  iter = c(3,3,3,3,3,3),
  start_lower = start_vals - 10,
  start_upper = start_vals + 10,
  lower = get_lower_lims(tpc_data2$temp_bin, tpc_data2$mean_rate,
                         model_name = "sharpeschoolfull_1981"),
  upper = get_upper_lims(tpc_data2$temp_bin, tpc_data2$mean_rate,
                         model_name = "sharpeschoolfull_1981"),
  supp_errors = "Y"
)
#Check the fit 
summary(mod)
#generate predictions for plotting 
preds <- data.frame(
  Temp = seq(min(tpc_data$Temp), max(tpc_data$Temp), length.out = 100)
)
preds <- broom::augment(mod, newdata = preds)

#plot TPC 
ggplot() +
  geom_point(data = tpc_data, aes(Temp, rate_mass)) +
  geom_line(data = preds, aes(Temp, .fitted), color = "blue") +
  theme_bw()

summary(tpc_data$rate_mass)

summary(tpc_data)
summary(mod)

#RESPR now we can calculate oxygen consumption rates 
#urchin_rates <- calc_rate.int(FRresp,
             # starts = c(1:600 17401, 34801, 52201,69601, 87001, #104401, 121801)) 
#summary(urchin_rates)
#plot(urchin_rates)
```

### Starting Over Again 

```{r}

# change in oxygen
delta_O2 <- O2_start - O2_end

# metabolic rate
MR <- (Oairsat2 / timepassed) * ChV_ml

freaky <- FR1_8 %>% #select for O ar saturation, chamber, temperautre, seconds
  select(Oairsat, Temp,chamber, time_parsed, seconds) %>% 
  #add in mass for each urchin
  mutate( 
    masskg = case_when(
      chamber == 1 ~ 20.58/1000,
      chamber == 2 ~ 13.08/1000,
      chamber == 3 ~ 20.54/1000,
      chamber == 4 ~ 1/10000,
      chamber == 5 ~ 13.34/1000, 
      chamber == 6 ~ 18.60/1000, 
      chamber == 7 ~ 17.07/1000, 
      chamber == 8 ~ 1/10000
    )
  ) %>% #Then we need to add in chamber volume 
   mutate( 
    ChV_ml = case_when(
      chamber == 1 ~ 743.92,
      chamber == 2 ~ 751.42,
      chamber == 3 ~ 743.96,
      chamber == 4 ~ 763.50,
      chamber == 5 ~ 751.16, 
      chamber == 6 ~ 745.90, 
      chamber == 7 ~ 747.43, 
      chamber == 8 ~ 763.50
    )
  ) %>% #add in a blank 1 & blank 2 column 
  mutate(
    blank = case_when(
      chamber <= 4 ~ 1,
      chamber >= 5 ~ 2
    )
  ) %>% 
  mutate(timepassed = 1) %>% 
  #need to add a column for change in oxygen o2 end 
  mutate(Oairsat2 = c(NA, diff(Oairsat) / diff(seconds))) 
   
      
head(freaky) #check out the data to see if selected correctly 
tibble(freaky)

# change in oxygen
FKEY_oxy_chg  <- freaky %>%
  mutate(
    delta_O2 = Oairsat2,
    metabolic_rate = ((delta_O2 / masskg * timepassed) 
  )

```

### calculate background rate 

```{r}
#this is where we need to adjust for the rate 
#the we need to group everything into temo_bin and avergae it for each temperature step
#going to group by blank > chamber > time 

#Get the blank rates 
blank1_rate <- FKEY_oxy_chg %>%
  filter(chamber == 4) %>%
  pull(Oairsat2) %>%
  first()

blank2_rate <- FKEY_oxy_chg %>%
  filter(chamber == 8) %>%
  pull(Oairsat2) %>%
  first()

#Create the blank_rate column 
FKEY_adj <- FKEY_oxy_chg %>%
  mutate(
    blank_rate = case_when(
      chamber %in% 1:3 ~ blank1_rate,
      chamber %in% 5:7 ~ blank2_rate,
      TRUE ~ NA_real_
    )
  )

#Minus the MR of the background 
FKEY_adj <- FKEY_adj %>% 
  mutate(O2.sec.corr = Oairsat2 - blank_rate) %>% 
  filter(!chamber == "4" | !chamber == "8") 
################################################################################
#lets try grouping it here 
FKEY_adj_group <- FKEY_adj %>% 
   filter(is.finite(O2.sec.corr)) %>% 
  mutate(Temp = round(Temp, 1)) %>%
  group_by(Temp, chamber, masskg) %>%
  summarise(O2.sec.corr.mean = mean(O2.sec.corr)) 

#find the mml cm2 hr
FKEY_adj_group_hr <- FKEY_adj_group %>% 
  mutate(umol.cm2.hr = (O2.sec.corr.mean*3600)/masskg
  ) %>% 
  filter(umol.cm2.hr > 0)

#log the rates
FKEY_adj_group_log <- FKEY_adj_group_hr %>% 
mutate(rate.in = log(umol.cm2.hr+0.1))

FKEY_adj_group_clean <- FKEY_adj_group_log %>% 
  filter(is.finite(rate.in))
#########################################################################################

#find the mml cm2 hr
FKEY_adj_hr <- FKEY_adj %>% 
  mutate(umol.cm2.hr = (O2.sec.corr*3600)/masskg
  )

#remove all na rows
FKEY_hr_na<- FKEY_adj_hr %>% 
  filter(is.finite(blank_rate))

#group it by temp 
FKEY_grouped <- FKEY_hr_na %>%
  group_by(Temp) %>%
  summarise(metabolic_rate = mean(umol.cm2.hr), .groups = "drop") %>% 
  filter(is.finite(metabolic_rate))

#log the rates
FKEY_log <- FKEY_hr_na %>% 
mutate(rate.in = log(umol.cm2.hr+0.1))
#filter out NAs
FKEY_log_clean <- FKEY_log %>% 
  filter(is.finite(rate.in))
```

### Now we can calculate MR

```{r}
#recalculate MR
FKEY_MR  <- FKEY_adj %>%
  mutate(
    delta_O2 = Oairsat2,
    metabolic_rate = (O2.sec.corr / masskg * timepassed) 
  )

#now lets group it by temperature 
FKEY_avg <- FKEY_MR %>%
  group_by(Temp) %>%
  summarise(metabolic_rate = mean(metabolic_rate), .groups = "drop")


#account for mass
#FKEY_mass <- FKEY_oxy_chg %>%
 # mutate(
  #  mass_specific_MR = metabolic_rate / masskg
  #)

#need the slope of oxygen decline 
#model <- lm(Oairsat ~ seconds, data = FKEY_mass)

#slope <- coef(model)[2]

#add MR as a column
#FKEY_MR <- FKEY_mass %>% 
 # mutate(metabolic_rate = slope * ChV_ml
 # )
#head(FKEY_MR)
```

#### Quick Visualizations

```{r}
MRplot <- FKEY_grouped %>% 
  ggplot(aes(x= Temp, y = metabolic_rate)) +
  geom_point()
MRplot

plot(FKEY_adj_group_log $Temp,
     FKEY_adj_group_log $rate.in,
     pch = 16,
     xlab = "Temperature (°C)",
     ylab = "Metabolic Rate",
     main = "TPC diagnostic plot")

plot(FKEY_adj_group_hr$Temp,
     log(FKEY_adj_group_hr$umol.cm2.hr),
     pch = 16,
     xlab = "Temperature",
     ylab = "log(Metabolic Rate)")
```

### 

### Blanks Data 

```{r eva; = FALSE}
#make a new data frame with only the blanks chamber
blank_data <- FKEY_MR %>%
  filter(chamber == 4 | chamber == 8)

#adjust for background rate 
blank_model <- lm(Oairsat ~ seconds, data = blank_data)
blank_slope <- coef(blank_model)[2]



```

### Model

```{r}
#remove Nas
FKEY_MR <- FKEY_MR %>%
  filter(is.finite(metabolic_rate), is.finite(Temp))

#fill in missing values
start_vals["el"] <- 1
#start_vals["eh"] <- 1

#recalculate starting values
start_vals <- get_start_vals(
  FKEY_MR$Temp,
  FKEY_MR$metabolic_rate,
  model_name = "sharpeschoolfull_1981"
)

#check vals
start_vals
is.finite(start_vals)

#check for negative or zero rates
summary(FKEY_MR$metabolic_rate)

#fit the model 
mod <- nls_multstart(
  metabolic_rate ~ sharpeschoolfull_1981(
    temp = Temp,
    r_tref,
    e,
    el,
    tl,
    eh,
    th,
    tref = 20
  ),
  data = FKEY_MR,
  iter = c(3,3,3,3,3,3),
  start_lower = start_vals - 10,
  start_upper = start_vals + 10,
  lower = get_lower_lims(FKEY_MR$Temp, FKEY_MR$metabolic_rate,
                         model_name = "sharpeschoolfull_1981"),
  upper = get_upper_lims(FKEY_MR$Temp, FKEY_MR$metabolic_rate,
                         model_name = "sharpeschoolfull_1981"),
  supp_errors = "Y"
)

#Check the fit 
summary(mod)
#generate predictions for plotting 
preds <- data.frame(
  Temp = seq(min(tpc_data$Temp), max(tpc_data$Temp), length.out = 100)
)
preds <- broom::augment(mod, newdata = preds)

#plot TPC 
ggplot() +
  geom_point(data = tpc_data, aes(Temp, rate_mass)) +
  geom_line(data = preds, aes(Temp, .fitted), color = "blue") +
  theme_bw()

summary(tpc_data$rate_mass)

summary(tpc_data)
summary(mod)
```
